---
title: "COVID-19 Dashboard"
output: 
  flexdashboard::flex_dashboard:
    theme: flatly
    orientation: rows
    social: menu
    source_code: embed
    vertical_layout: fill
---

<style>                     
.navbar {
  background-color:black;
  border-color:black;
}
</style>      

```{r, include=FALSE, warnings=FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

```{r dependencies, include=FALSE, warnings=FALSE, message=FALSE}
#------------------ Packages ------------------
## For Dashboard
library(flexdashboard)

## Data cleaning 
library(data.table)
library(dplyr)
library(tidyr)
library(magrittr)
library(purrr)
library(forcats)

## printing 
library(glue)

## Maps and Plots
library(plotly)
library(leaflet)
library(leafpop)
library(RColorBrewer)

# Dates
library(lubridate)

## For displaying tables
library(DT)

theme_set(theme_classic())
```

```{r cases_data}
#------------------ Data ------------------
url <- "https://raw.githubusercontent.com/RamiKrispin/coronavirus/master/csv/coronavirus.csv"
df <- fread(url) %>%
    mutate(
        country = trimws(country),
        country = factor(country, levels = unique(country))
    )

# for time series plot
df_daily <- df %>%
    group_by(date, type) %>%
    summarise(total = sum(cases, na.rm = T), .groups = "drop") %>%
    pivot_wider(names_from = type,
                values_from = total) %>%
    arrange(date) %>%
    ungroup() %>%
    mutate(active =  confirmed - death - recovered) %>%
    mutate(
        confirmed_cum = cumsum(confirmed),
        death_cum = cumsum(death),
        recovered_cum = cumsum(recovered),
        active_cum = cumsum(active)
    )

df_country <- df %>%
    group_by(country, type) %>%
    summarise(total = sum(cases), .groups = "drop") %>%
    pivot_wider(names_from = type, values_from = total) %>%
    mutate(active = confirmed - death - recovered) %>%
    pivot_longer(cols = -country,
                 names_to = "type",
                 values_to = "total")

# for global statistics
df_world <- df_country %>%
    group_by(type) %>%
    summarise(total = sum(total), .groups = "drop") %>%
    pivot_wider(names_from = type, values_from = total)

# for leaflet map
df_map <- df %>%
    filter(cases > 0) %>%
    group_by(country, province, lat, long, type) %>%
    summarise(cases = sum(cases), .groups = "drop") %>%
    mutate(log_cases = 2 * log(cases)) %>%
    ungroup()

df_map.split <- df_map %>%
    split(df_map$type)
```


```{r vax_data}
# Vaccination Data

url2 <- "https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv"

df_vax <- fread(url2) %>% 
  replace(is.na(.), 0) %>% 
  mutate(location = factor(location, levels = unique(location))) %>% 
  as_tibble()

# glimpse(df_vacc)

## daily vaccinations
df_vax_daily <- df_vax %>% 
  select(date, location, daily_vaccinations) %>% 
  group_by(location, date) %>% 
  summarise(total = sum(daily_vaccinations), .groups="drop") %>%
  filter(location != "World") %>% 
  arrange(-total) %>%
  pivot_wider(names_from = date, values_from = total) %>%
  # replace(is.na(.), 0) %>% 
  # mutate(across(where(is.numeric), log)) %>% 
  slice(1:20) %>% 
  pivot_longer(cols = -location, 
               names_to = "date", 
               values_to = "total") %>% 
  mutate(date = ymd(date)) 


# world statistics
df_world_vax <- df_vax %>% 
  select(location,
         total_vaccinations,
         people_vaccinated,
         people_fully_vaccinated,
  ) %>% 
  filter(location == "World") %>% 
  summarise(total_vax = max(total_vaccinations),
            ppl_once = max(people_vaccinated),
            fully_vax = max(people_fully_vaccinated)) %>% 
  mutate(across(where(is.numeric), scales::comma))

# total vaccines administered
df_vax_total <- df_vax %>% 
  select(location, iso_code, total_vaccinations) %>% 
  group_by(location) %>% 
  filter(total_vaccinations == max(total_vaccinations), location != "World") 
```


Overview
=======================================================================

Row
-----------------------------------------------------------------------

### confirmed {.value-box}

```{r}
confirmed <- format(df_world$confirmed, big.mark = ",")
valueBox(value = confirmed, 
         caption = "Total Confirmed Cases", 
         color = "black")
```


### active {.value-box}

```{r}
active <- format(df_world$active, big.mark = ",")
active_perc <- round(100 * df_world$active / df_world$confirmed, 1)

valueBox(
    value = glue("{active} ({active_perc}%)"),
    caption = "Active Cases",
    color = "blue"
)
```

### recovered {.value-box}

```{r}
recovered <- format(df_world$recovered, big.mark = ",")
recovered_perc <-
    round(100 * df_world$recovered / df_world$confirmed, 1)

valueBox(
    value = glue("{recovered} ({recovered_perc}%)"),
    caption = "Recovered Cases",
    color = "green"
)
```

### death {.value-box}

```{r}
death <- format(df_world$death, big.mark = ",")
death_perc <- round(100 * df_world$death / df_world$confirmed, 1)

valueBox(
    value = glue("{death} ({death_perc}%)"),
    caption = "Death Cases",
    color = "red"
)
```


Row {.tabset}
-----------------------------------------------------------------------

### Map of Cases

```{r}
# Color pallete
pal <- colorFactor(c("grey", "red", "green"),
                   domain = c("confirmed", "death", "recovered"))

map_object <- leaflet() %>%
    addProviderTiles(providers$CartoDB.DarkMatter) %>% 
    setView(41.405559, 0.247590, zoom=2)

names(df_map.split) %>%
  walk(function(df) {
    map_object <<- map_object %>%
      addCircleMarkers(data=df_map.split[[df]],
                 lng=~long, lat=~lat,
                 color = ~pal(type),
                 stroke = FALSE,
                 fillOpacity = 0.5,
                 radius = ~log_cases,
                 popup =  popupTable(df_map.split[[df]],
                                              feature.id = FALSE,
                                              row.numbers = FALSE,
                                              zcol=c("type","cases","country","province")),
                 group = df,
                 labelOptions = labelOptions(noHide = F,
                                             direction = 'auto'))
  })

map_object %>%
  addLayersControl(
    overlayGroups = names(df_map.split),
    options = layersControlOptions(collapsed = FALSE)) %>% 
    hideGroup(c("confirmed", "recovered"))
```


Daily 
=======================================================================

Column {data-width=400}
-------------------------------------

### Daily Cumulative Cases
    
```{r}
fig <- plot_ly(
    data = df_daily,
    x = ~ date,
    y = ~ active_cum,
    name = "Active",
    fillcolor = "blue",
    type = "scatter",
    mode = "none",
    stackgroup = "one") %>%
    add_trace(y = ~ recovered_cum,
              name = "Recovered",
              fillcolor = "green") %>%
    add_trace(y = ~ death_cum,
              name = "Death",
              fillcolor = "red") %>%
    layout(
        title = "",
        yaxis = list(title = "Cumulative Number of Cases"),
        xaxis = list(title = "Date",
                     type = "date"),
        legend = list(x = 0.1, y = 0.9),
        hovermode = "compare"
    )
fig
```


### Histogram of Daily Cases

```{r}
g <- ggplot(data=df_daily, aes(date, confirmed)) +
    geom_histogram(stat="identity") +
    scale_y_continuous(labels = scales::label_number_si())

ggplotly(g)
```


Top 20 Countries
=======================================================================

Column {.tabset}
-------------------------------------

```{r}
top_20 <- df_country %>% 
    group_by(country, type) %>% 
    pivot_wider(names_from = type, values_from=total) %>% 
    arrange(-confirmed) %>% 
    ungroup() %>% 
    top_n(20)

# top_20  %>%
#     kbl() %>%
#     kable_styling(
#         bootstrap_options = c("striped", "hover", "condensed", "responsive"),
#         position = "left",
#         full_width = F
#     ) %>% 
#     scroll_box(width = "100%", height = "100%")

# define 20 colors with colorRampPalette
nb.cols <- 20
mycolors <- colorRampPalette(brewer.pal(10, "RdYlGn"))(nb.cols)
```

### Confirmed Cases
```{r}
confirmed <- top_20 %>% 
    mutate(country = fct_reorder(country, confirmed)) %>% 
    ggplot(top_20, mapping = aes(x=country, y=confirmed, fill=country)) + 
    geom_bar(stat='identity') +
    theme(legend.position = 'none') +
    scale_fill_manual(values = rev(mycolors)) +
    scale_y_continuous(labels = scales::label_number_si()) +
    labs(y="confirmed cases", x="") +
    coord_flip() +
    ggtitle("Top 20 countries for Confirmed Cases")

ggplotly(confirmed, tooltip="confirmed")
```

### Death Cases
```{r}
death <- top_20 %>% 
    mutate(country = fct_reorder(country, death)) %>% 
    ggplot(top_20, mapping = aes(x=country, y=death, fill=country)) + 
    geom_bar(stat='identity') +
    theme(legend.position = 'none') +
    scale_fill_manual(values = rev(mycolors)) +
    scale_y_continuous(labels = scales::label_number_si()) +
    labs(y="Death Cases", x="") +
    coord_flip() +
    ggtitle("Top 20 countries for Death Cases")

ggplotly(death, tooltip="death")  
```

### Active Cases

```{r}
active<- top_20 %>% 
    mutate(country = fct_reorder(country, active)) %>% 
    ggplot(top_20, mapping = aes(x=country, y=active, fill=country)) + 
    geom_bar(stat='identity') +
    theme(legend.position = 'none') +
    scale_fill_manual(values = rev(mycolors)) +
    scale_y_continuous(labels = scales::label_number_si()) +
    labs(y="Active Cases", x="") +
    coord_flip() +
    ggtitle("Top 20 countries for Active Cases")

ggplotly(active, tooltip="active")
```

### Cases Data

```{r}
df %>%
  group_by(country, type) %>%
  summarise(total = sum(cases), .groups = "drop") %>%
  pivot_wider(names_from = type, values_from = total) %>%
  mutate(active = confirmed - death - recovered) %>%
  datatable(options = list(
    bPaginate = FALSE,
    scrollX = TRUE,
    columnDefs = list(list(className = 'dt-center', targets = "_all"))
  ),
  rownames=FALSE,
  caption = 'Cases Data',
  style = 'bootstrap', 
  class = 'cell-border stripe')
```

Vaccination 
=======================================================================

Row
-----------------------------------------------------------------------

### Total Doses {.value-box}

```{r}
valueBox(value = df_world_vax$total_vax,
         caption = "Total doses administered",
         color = "#228B22")
```


### At least one {.value-box}

```{r}
valueBox(value = df_world_vax$ppl_once,
         caption = "No. of people vaccinated (at least one dose)",
         color = "#228B22")
```


### People Fully vaccinated {.value-box}

```{r}
valueBox(value = df_world_vax$fully_vax, 
         caption = "No. of people fully vaccinated", 
         color = "#228B22")
```


Column {.tabset}
-----------------------------------------------------------------------

```{r}
# scales::number(total, big.mark = "", accuracy = 1))
# vax_daily <-
#   ggplot(df_vax_daily, aes(x = date, y = total, color = location, )) +
#   geom_line() +
#   theme(legend.position = 'none') +
#   scale_colour_manual(values = mycolors) +
#   scale_x_date(date_labels = "%b %d") +
#   scale_y_continuous(labels = scales::label_number()) +
#   ggtitle("Daily COVID-19 vaccine doses administered")
# 
# ggplotly(vax_daily)
```

### Map

```{r}
# figure out how to use iso code to create leaflet maps
# https://www.r-graph-gallery.com/183-choropleth-map-with-leaflet.html

# Plotly choroplet map
# https://plotly.com/r/choropleth-maps/
# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)
# add trace
vax_map <- plot_geo(df_vax_total)
vax_map <- vax_map %>% add_trace(
  z = ~ total_vaccinations,
  color = ~ total_vaccinations,
  colors = 'Greens',
  text = ~ location,
  locations = ~ iso_code,
  marker = list(line = l)
)

# add title to colorbar
vax_map <- vax_map %>% colorbar(title = 'Total vaccinations (M)')

# add title and subtitle
vax_map <- vax_map %>%
  layout(title = 'Total COVID-19 Vaccination Doses Administered<br>Source:<a href="https://github.com/owid/covid-19-data/tree/master/public/data/vaccinations">Our World in Data</a>', geo = g)

vax_map
```

### Top 20

```{r}
# for limits of graph
max = arrange(df_vax_total, -total_vaccinations)$total_vaccinations[1]

vax_total <- df_vax_total %>%
  arrange(-total_vaccinations) %>% 
  ungroup() %>% 
  slice(1:20) %>% 
  mutate(location = fct_reorder(location, total_vaccinations)) %>%
  ggplot(aes(
    x = total_vaccinations,
    y = location,
    fill = location,
    label = scales::label_number_si(accuracy = 0.01)(total_vaccinations)
  )) + 
  geom_bar(stat = "identity") +
  geom_text(color="grey50", size = 4) +
  theme(legend.position = 'none') +
  scale_fill_manual(values = mycolors) +
  scale_x_continuous(labels = scales::label_number_si(), limits = c(0, max+20e6)) +
  labs(x="total vaccinations", y="")

ggplotly(vax_total) %>% 
  style(hoverinfo = "none", textposition = "right")
```

### Vaccination data
```{r}
df_vax_total %>% 
  select(location, total_vaccinations) %>% 
  datatable(options = list(
    bPaginate = FALSE,
    scrollX = TRUE,
    columnDefs = list(list(className = 'dt-center', targets = "_all"))
  ),
  rownames=FALSE,
  caption = 'Vaccination Data',
  style = 'bootstrap', 
  class = 'cell-border stripe')
```

About
=======================================================================

This dashboard provides an overview of the COVID-19 (2019-nCoV) epidemic. This dashboard was built with R using Rmarkdown and the Flexdashboard package. The Github Repository for this dashboard available [here](https://github.com/benthecoder/covid19-flexdashboard)

**Data**

The source of data for this dashboard is from the R package [coronavirus](https://github.com/RamiKrispin/coronavirus). With the raw data pulled from the [COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University](https://github.com/CSSEGISandData/COVID-19)

The source of vaccination data can be found at the Our World in Data [vaccination github repo](https://github.com/owid/covid-19-data/tree/master/public/data/vaccinations)

**Packages**

* Dashboard interface - `flexdashboard`
* Visualization - `plotly`, `RColorBrewer`, `ggplot2`
* Data Import - `data.table`
* Data manipulation - `dplyr`, `tidyr`, `purrr`, `forcats`
* Mapping - `leaflet`, `leafpop` 
* Strings - `glue`
* Data Source - `coronavirus`
* Data display - `DT`
* Dates - `lubridate`
* Scales - `scales`

**Update**

* This dashboard was last updated at **`r paste0(format(Sys.time(), "%A, %B %d %Y at %X"), " (", Sys.timezone(), ")")`**
* The latest date from the data is **`r max(df$date)`**
* The latest date from the vaccination data is **`r max(df_vax$date)`** 
