---
title: "COVID-19 Dashboard"
author: "Benedict Neo"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: embed
    vertical_layout: fill
---

<style>                     
.navbar {
  background-color:black;
  border-color:black;
}
</style>      

```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
#------------------ Packages ------------------
## For Dashboard
library(flexdashboard)

## Data cleaning 
library(dplyr)
library(tidyr)
library(magrittr)
library(purrr)
library(forcats)

## printing 
library(glue)

## Maps and Plots
library(plotly)
library(leaflet)
library(leafpop)
library(RColorBrewer)

library(coronavirus)
data(coronavirus)
# View(coronavirus)

#------------------ Data ------------------
df <- coronavirus %>%
    mutate(
        country = trimws(country),
        country = factor(country, levels = unique(country))
    )

# for time series plot
df_daily <- coronavirus %>%
    group_by(date, type) %>%
    summarise(total = sum(cases, na.rm = T), .groups = "drop") %>%
    pivot_wider(names_from = type,
                values_from = total) %>%
    arrange(date) %>%
    ungroup() %>%
    mutate(active =  confirmed - death - recovered) %>%
    mutate(
        confirmed_cum = cumsum(confirmed),
        death_cum = cumsum(death),
        recovered_cum = cumsum(recovered),
        active_cum = cumsum(active)
    )

df_tree <- df %>%
    group_by(country, type) %>%
    summarise(total = sum(cases), .groups = "drop") %>%
    pivot_wider(names_from = type, values_from = total) %>%
    mutate(active = confirmed - death - recovered) %>%
    pivot_longer(cols = -country,
                 names_to = "type",
                 values_to = "total")

# for global statistics
df_world <- df_tree %>%
    group_by(type) %>%
    summarise(total = sum(total), .groups = "drop") %>%
    pivot_wider(names_from = type, values_from = total)

# for leaflet map
df_map <- df %>%
    filter(cases > 0) %>%
    group_by(country, province, lat, long, type) %>%
    summarise(cases = sum(cases), .groups = "drop") %>%
    mutate(log_cases = 2 * log(cases)) %>%
    ungroup()

df_map.split <- df_map %>%
    split(df_map$type)
```


Overview
=======================================================================

Row
-----------------------------------------------------------------------

### confirmed {.value-box}

```{r}
confirmed <- format(df_world$confirmed, big.mark = ",")
valueBox(value = confirmed, 
         caption = "Total Confirmed Cases", 
         color = "black")
```


### active {.value-box}

```{r}
active <- format(df_world$active, big.mark = ",")
active_perc <- round(100 * df_world$active / df_world$confirmed, 1)

valueBox(
    value = glue("{active} ({active_perc}%)"),
    caption = "Active Cases",
    color = "blue"
)
```

### recovered {.value-box}

```{r}
recovered <- format(df_world$recovered, big.mark = ",")
recovered_perc <-
    round(100 * df_world$recovered / df_world$confirmed, 1)

valueBox(
    value = glue("{recovered} ({recovered_perc}%)"),
    caption = "Recovered Cases",
    color = "green"
)
```

### death {.value-box}

```{r}
death <- format(df_world$death, big.mark = ",")
death_perc <- round(100 * df_world$death / df_world$confirmed, 1)

valueBox(
    value = glue("{death} ({death_perc}%)"),
    caption = "Death Cases",
    color = "red"
)
```


Row {.tabset}
-----------------------------------------------------------------------

### Map of Cases

```{r}
# Color pallete
pal <- colorFactor(c("grey", "red", "green"),
                   domain = c("confirmed", "death", "recovered"))

map_object <- leaflet() %>%
    addProviderTiles(providers$CartoDB.DarkMatter) %>% 
    setView(41.405559, 0.247590, zoom=2)

names(df_map.split) %>%
  walk(function(df) {
    map_object <<- map_object %>%
      addCircleMarkers(data=df_map.split[[df]],
                 lng=~long, lat=~lat,
                 color = ~pal(type),
                 stroke = FALSE,
                 fillOpacity = 0.5,
                 radius = ~log_cases,
                 popup =  popupTable(df_map.split[[df]],
                                              feature.id = FALSE,
                                              row.numbers = FALSE,
                                              zcol=c("type","cases","country","province")),
                 group = df,
                 labelOptions = labelOptions(noHide = F,
                                             direction = 'auto'))
  })

map_object %>%
  addLayersControl(
    overlayGroups = names(df_map.split),
    options = layersControlOptions(collapsed = FALSE)) %>% 
    hideGroup(c("confirmed", "recovered"))
```


Daily Cumulative Cases 
=======================================================================

Column {data-width=400}
-------------------------------------

### Time Series
    
```{r}
fig <- plot_ly(
    data = df_daily,
    x = ~ date,
    y = ~ active_cum,
    name = "Active",
    fillcolor = "blue",
    type = "scatter",
    mode = "none",
    stackgroup = "one") %>%
    add_trace(y = ~ recovered_cum,
              name = "Recovered",
              fillcolor = "green") %>%
    add_trace(y = ~ death_cum,
              name = "Death",
              fillcolor = "red") %>%
    layout(
        title = "",
        yaxis = list(title = "Cumulative Number of Cases"),
        xaxis = list(title = "Date",
                     type = "date"),
        legend = list(x = 0.1, y = 0.9),
        hovermode = "compare"
    )
fig
```


### Histogram of Daily Cases

```{r}
g <- ggplot(data=df_daily, aes(date, confirmed), color="black") +
    geom_histogram(stat="identity") +
    theme_classic() +
    scale_y_continuous(labels = scales::label_number_si())

ggplotly(g)
```


Top 20 Countries
=======================================================================

Column {.tabset}
-------------------------------------

```{r}
top_20 <- df_tree %>% 
    group_by(country, type) %>% 
    pivot_wider(names_from = type, values_from=total) %>% 
    arrange(-confirmed) %>% 
    ungroup() %>% 
    top_n(20)

# top_20  %>%
#     kbl() %>%
#     kable_styling(
#         bootstrap_options = c("striped", "hover", "condensed", "responsive"),
#         position = "left",
#         full_width = F
#     ) %>% 
#     scroll_box(width = "100%", height = "100%")

nb.cols <- 20
mycolors <- colorRampPalette(brewer.pal(10, "RdYlGn"))(nb.cols)
```

### Confirmed Cases
```{r}
confirmed <- top_20 %>% 
    mutate(country = fct_reorder(country, confirmed)) %>% 
    ggplot(top_20, mapping = aes(x=country, y=confirmed, fill=country)) + 
    geom_bar(stat='identity') +
    theme_classic() +
    theme(legend.position = 'none') +
    scale_fill_manual(values = rev(mycolors)) +
    scale_y_continuous(labels = scales::label_number_si()) +
    coord_flip() +
    labs(x="Confirmed Cases", title="Top 20 countries for Confirmed Cases")

ggplotly(confirmed)
```

### Death Cases
```{r}
death <- top_20 %>% 
    mutate(country = fct_reorder(country, death)) %>% 
    ggplot(top_20, mapping = aes(x=country, y=death, fill=country)) + 
    geom_bar(stat='identity') +
    theme_classic() +
    theme(legend.position = 'none') +
    scale_fill_manual(values = rev(mycolors)) +
    scale_y_continuous(labels = scales::label_number_si()) +
    coord_flip() +
    labs(x="Death Cases", title="Top 20 countries for Death Cases")

ggplotly(death)
```

### Active Cases

```{r}
active<- top_20 %>% 
    mutate(country = fct_reorder(country, active)) %>% 
    ggplot(top_20, mapping = aes(x=country, y=active, fill=country)) + 
    geom_bar(stat='identity') +
    theme_classic() +
    theme(legend.position = 'none') +
    scale_fill_manual(values = rev(mycolors)) +
    scale_y_continuous(labels = scales::label_number_si()) +
    coord_flip() +
    labs(x="Active Cases", title="Top 20 countries for Active Cases")

ggplotly(active)
```


About
=======================================================================

**The COVID-19 Dashboard**

This dashboard provides an overview of the COVID-19 (2019-nCoV) epidemic. This dashboard was built with R using Rmarkdown and the Flexdashboard package. The Github Repository for this dashboard available [here](https://github.com/benthecoder/covid19-flexdashboard)

**Data**

The source of data for this dashboard is from the R package [coronavirus](https://github.com/RamiKrispin/coronavirus). With the raw data pulled from the [COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University](https://github.com/CSSEGISandData/COVID-19)

**Packages**

* Dashboard interface - `flexdashboard`
* Visualization - `plotly`, `RColorBrewer`
* Data manipulation - `dplyr`, `tidyr`, `purrr`, `forcats`
* Mapping - `leaflet`, `leafpop` 
* Strings - `glue`
* Data Source - `coronavirus`

This dashboard was last updated at `r max(df$date)`
